---
title: "Mood tracker"
author: "Ava Gaines"
date: "2026-04-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(readxl)
library(ggplot2)
```

## Load in and display data frame
```{r}
MoodTracker <- read_excel("~/DATA 211/MoodTracker.xlsx") %>%
  select(Observation, DayOfWeek, Outside, Mood)

print(MoodTracker, n=21)
```
## Generate summary statistics
```{r}
MoodTracker %>%
  group_by(Outside) %>%
  summarise(
    n         = n(),
    mean      = mean(Mood),
    median    = median(Mood),
    sd        = sd(Mood),
    min       = min(Mood),
    max       = max(Mood)
  )
```

## Hypotheses
$\mu_o$: average mood on days I went outside

$\mu_i$: average mood on days I stayed inside

$H_0: \mu_o = \mu_i$ vs $H_a: \mu_o > \mu_i$

## Run two-sample T-test
```{r}
outside <- MoodTracker$Mood[MoodTracker$Outside == "Yes"]
inside  <- MoodTracker$Mood[MoodTracker$Outside == "No"]

t.test(outside, inside, alternative = "greater" , var.equal = FALSE)  
```
Because p-value is greater than $\alpha = 0.05$, we fail to reject the null hypothesis. In this context, that means we do not have sufficient evidence to conclude that my mood on days I went outside is significantly higher than my mood on days I stayed inside.

## Create boxplot using data
```{r}
ggplot(MoodTracker, aes(x = Outside, y = Mood, fill = Outside)) +
  geom_boxplot() +
  scale_y_continuous(limits = c(1, 5), breaks = 1:5) +
  labs(
    title = "Mood Score by Outside Activity",
    x = "Went Outside",
    y = "Mood Score"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

